Fig3

### 2020-05-06 analysis/figure-2.Rmd
better PRL fig, now fig 3 not 4

b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.1,1.1, 1,1, 1,1))

>>>>

b <- plot_grid(b1,b2,b3, ncol = 1, rel_heights = c(1.1,1,1))
c <- plot_grid(c1,c2,c3,c4,c5,c6, ncol = 2, rel_heights = c(1.1,1,1), rel_widths = c(1.1,1))
d <- plot_grid(d1,d2,d3,d4,d5,d6, ncol = 1, rel_heights = c(1.2,1, 1,1, 1,1))

### 2020-05-06 analysis/figure-3.Rmd
better PRL fig, now fig 3 not 4

---
title: "Fig3: Correlations in gene expression"
author: "Rayna M Harris"
date: "3/26/2020"
output: md_document
---

# Correlations in gene expression

```{r setup}
library(tidyverse)
library(cowplot)
library(corrr)
library(ggpubr)

@@

knitr::opts_chunk$set(echo = TRUE, message = F, fig.path = "../figures/")
```

## data wrangle

```{r data}

# load `candidatevsd` with `source("../R/wrangledata.R")`
head(candidatevsd)

curleychampagnegenes <- curleychampagnegenes %>% pull(gene)

# widen for correlations

c1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
c2 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
c3 <- makecorrdf("female", "gonad", curleychampagnegenes)

c4 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
c5 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
c6 <- makecorrdf("male", "gonad", curleychampagnegenes)

candidatevsdwide <- candidatevsd  %>%
    pivot_wider(names_from = gene, values_from = counts) 
head(candidatevsdwide)

FH <- subsetcandidatevsdwide("female", "hypothalamus")
FP <- subsetcandidatevsdwide("female", "pituitary")
FG <- subsetcandidatevsdwide("female", "gonad")
MH <- subsetcandidatevsdwide("male", "hypothalamus")
MP <- subsetcandidatevsdwide("male", "pituitary")
MG <- subsetcandidatevsdwide("male", "gonad") 
```

### figure

```{r fig3, fig.width = 7.25, fig.height = 7.25}

p1 <- plotcorrplot(c1, "FH") + theme(legend.position = "none", axis.text.x = element_blank()) + labs(subtitle = "Hypothalamus")
p2 <- plotcorrplot(c2, "FP") + theme(legend.position = "none", axis.text.x = element_blank()) + 
  labs(subtitle = "Pituitary") 
p3 <- plotcorrplot(c3, "FG") + theme(legend.position = "none") + labs(subtitle = "Gonad")

p4 <- plotcorrplot(c4, "MH") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank()) + labs(subtitle = " ")
p5 <- plotcorrplot(c5, "MP") + theme(legend.position = "none", axis.text.y = element_blank(), axis.text.x = element_blank())  +
  labs(subtitle = " ") 
p6 <- plotcorrplot(c6, " ") + theme(legend.position = "none", axis.text.y = element_blank()) + labs(subtitle = " ")

a <- plot_grid(p1 + labs(title = "Female"),
               p4 + labs(title = "Male"),
               p2, p5, p3, p6, nrow = 3, rel_widths = c(1.2,1),
               rel_heights = c(1,1,1.2))

scattercorrelations <- function(df, gene1, gene2, mylinecolor, myxlab, myylab){
  p <- ggplot(df, aes(x = gene1, y = gene2)) +
    labs(subtitle = " ", x = myxlab, y = myylab) + 
    scale_color_manual(values = allcolors) +
    geom_smooth(method = "lm",  color = mylinecolor) +
    geom_point(aes(color = treatment))  +
    theme_B3() +  
    theme(axis.title = element_text(face = "italic"), 
            legend.position = "none") +
    stat_cor( size = 2)
  return(p)
}

# "female" = "#969696", "male" = "#525252"

p7 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + labs(subtitle = "Hypothalamus")
p8 <- scattercorrelations(FP, FP$AVPR1A, FP$CRHR1, "#969696", "AVPR1A", "CRHR1")  + labs(subtitle = "Pituitary")
p9 <- scattercorrelations(FG, FG$ESR2, FG$CRHBP, "#969696", "ESR2", "CRHBP")   + labs(subtitle = "Gonad")

p10 <- scattercorrelations(MH, MH$DRD1, MH$HTR2C, "#525252", "DRD1", "HTR2C") + theme(axis.title.y = element_blank())
p11 <- scattercorrelations(MP, MP$AVPR1A, MP$CRHR1, "#525252", "AVPR1A", "CRHR1") + theme(axis.title.y = element_blank())
p12 <- scattercorrelations(MG, MG$ESR2, MG$CRHBP, "#525252", "ESR2", "CRHBP")  + theme(axis.title.y = element_blank())

b <- plot_grid(p7 + labs(title = "Female"), 
               p10 + labs(title = "Male"), 
               p8, p11, p9, p12, nrow = 3, rel_heights = c(1,1,1.2))

ab <- plot_grid(a,b, rel_widths = c(1.5,1), labels = c("A", "B"), label_size = 8)

p13 <- plotcorrplot(c1, "FH") + theme(legend.position = "bottom")
p14 <- scattercorrelations(FH, FH$DRD1, FH$HTR2C, "#969696", "DRD1", "HTR2C") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

legend1 <- p13 %>% get_legend()
legend2 <- p14 %>% get_legend()

legends <- plot_grid(legend1, legend2, rel_widths = c(1.5,1))

plot_grid(ab, legends, nrow = 2, rel_heights = c(1,0.1))

@@

>>>>

---
title: "Fig3"
output: md_document
---

# Figure 4: All things prolactin

```{r setup}
library(tidyverse)
library(cowplot)
library(ggimage)
library(apaTables)
library(factoextra)
library(corrr)
library(ggpubr)
library(ggrepel)
library(ggsignif)

@@

knitr::opts_chunk$set(fig.path = '../figures/',message=F, warning=FALSE)
```

## PCA data

```{r PCA}
# pca
pca1f <- subsetmakepca("pituitary", charlevels, "female")	
pca2f <- makefvizdf("pituitary", charlevels, "female")

pca1m <- subsetmakepca("pituitary", charlevels, "male")	
pca2m <- makefvizdf("pituitary", charlevels, "male")

@@

# subset PRL vsd

```{r}
# `candidatevsd` loaded with `source("../R/wrangledata.R")`

PRLpit <- candidatevsd %>% filter(tissue == "pituitary", gene == "PRL") %>%
  mutate(hiloPRL = ifelse(counts >= 18, "hi", "lo"))  %>%
  drop_na()
PRLpit$hiloPRL <- factor(PRLpit$hiloPRL, levels = c("lo", "hi"))

PRLpit %>%
  group_by(sex) %>%
  summarize(median = median(counts))

PRLpit %>%
  group_by(sex, tissue, hiloPRL) %>%
  summarize(n = n())

PRLpitF <- PRLpit %>% filter( sex == "female" )
PRLpitM <- PRLpit %>% filter( sex == "male")

## Internal versus external hyotheses

```{r hypothesis}

DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG2 <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG2)

allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)

allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
                                                      "eggs vs. chicks" = "eggs_chicks"))
allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))

PRLDEGs <- allDEG2 %>%
  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",
         direction == "hi") %>%
  arrange(desc(comparison))
PRLDEGs

## genes WGCNA prl module

PRLgenes <- read_csv("../results/PRLmodule.csv") %>% pull(x)

candidatevsd <- getcandidatevsd(PRLgenes, "pituitary", sexlevels)

df <- candidatevsd %>%
    pivot_wider(names_from = gene, values_from = counts) %>%
    select(-sex,-tissue, -treatment, -samples) %>%
    correlate() %>%
  focus(PRL)  %>%
  arrange(rowname)
df

```{r fig3, fig.width=7.25, fig.height=7.25}

a1 <- plotpc12(pca1f, pca2f, pca1f$treatment, allcolors, "Female pituitary") 
a2 <- plotpc12(pca1m, pca2m, pca1m$treatment, allcolors, "Male pituitary")
a <- plot_grid(a1,a2, labels = c("A"),label_size = 8)

bcd1 <- png::readPNG("../figures/images/fig_fig3a.png")
bcd1 <- ggdraw() +  draw_image(bcd1, scale = 1)

b2 <- plotprolactin(PRLpitF, PRLpitF$counts, "PRL", "Female gene expression") 
c2 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2200)   + theme(axis.text.x = element_blank() )
d2 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2200) + theme(axis.text.x = element_blank())   
b3 <- plotprolactin(PRLpitM, PRLpitM$counts, "PRL", "Male gene expression") + theme(axis.text.x = element_text())
c3 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2200)  
d3 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2200)

bcd23 <- plot_grid(b2,c2,d2,b3,c3,d3, nrow = 2, rel_widths = c(2,1,1), rel_heights = c(1,1.1))

## top correlations
e <- df %>%
  arrange(desc(PRL)) %>% 
  head(10)  %>%
  mutate(PRLrounded = round(PRL, 2))%>% 
  ggplot(aes(x = reorder(rowname, PRL), y = PRL)) +
  geom_bar(stat = "identity") +
  theme_B3() +
  theme(axis.text.y = element_text(face = "italic")) +
  labs(x = " ", y = "Correlation with PRL", subtitle = "Top 10 co-regulated genes") +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0))

# most sig DEGs

f <-  PRLDEGs %>%
  arrange(desc(lfc)) %>%
  filter(sex == "female") %>% head(10) %>% 
  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
  geom_bar(stat = "identity", fill = "#969696") +
  theme_B3() +
  theme(axis.text.y = element_text(face = "italic")) +
  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 female DEGs") +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0))

g <-  PRLDEGs %>%
  arrange(desc(lfc)) %>%
  filter(sex == "male") %>% head(10) %>% 
  ggplot(aes(x = reorder(gene, lfc), y = lfc)) + 
  geom_bar(stat = "identity", fill = "#525252") +
  theme_B3() +
  theme(axis.text.y = element_text(face = "italic")) +
  labs(x = " ", y = "Log-fold change, lo vs. hi PRL", subtitle = "Top 10 male DEGs") +
  coord_flip() +
  scale_y_continuous(expand = c(0, 0))

efg <- plot_grid(e,f,g, nrow = 1, labels = c("E", "F", "G"), label_size = 8)

plot_grid(a, bcd1, bcd23, efg, nrow = 4, rel_heights = c(1,1,2,1.2))

```{r savefiles}
write.csv(PRLpit,"../results/PRLvsd.csv", row.names = F)
